---
title: "01_exp_data"
author: "Puvvula"
date: "2023-07-17"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, haven)

dat_loc <- "~/Documents/ope_ssis/data/"
res<- "~/Documents/ope_ssis/result/"
```

#outcome and covariates data [n=236]
```{r}
outcom_cov<- read_csv("~/Documents/ope_ssis/res_sing_imput/out_cov.csv")
```

#ope data - pre and post natal exposure data received from aimin on Dec 13, 2023
#imputed 32 observations with no SG values with visit specific median SG
```{r}
dat_1<- read_sas(paste0(dat_loc, "raw/dec_23_new_ver/ur_fr_sg_20210921.sas7bdat"))
dat_2<- read_sas(paste0(dat_loc, "raw/dec_23_new_ver/ur_fr_12y.sas7bdat"))

ope_dat<- bind_rows(dat_1, dat_2)

ope_exp_raw<- ope_dat |>
  clean_names() |>
  filter(!analyte_code %in% c("DBzP", "DCP", "TBBA", "BCPP", "iPPPP", "tBPPP"))|>
  group_by(visit) |>
  mutate(specific_gravity = if_else(is.na(specific_gravity) | specific_gravity == 1,  
                                    median(specific_gravity, na.rm = TRUE),  
                                    specific_gravity)) |>
  ungroup() |>
  mutate(visit = case_when(
    visit == "16W" ~ "prenatal_16W",
    visit == "26W" ~ "prenatal_26W",
    visit == "Birth" ~ "delivery",
    visit == "12M" ~ "postnatal_1_yr",
    visit == "24M" ~ "postnatal_2_yr",
    visit == "36M" ~ "postnatal_3_yr",
    visit == "60M" ~ "postnatal_5_yr",
    visit == "P3" ~ "postnatal_8_yr",
    visit == "P4" ~ "postnatal_12_yr",
    TRUE ~ as.character(visit)
  )) |>
  mutate(analyte_code = fct_recode(as.factor(analyte_code), 
                             bcep = "BCEtP", bdcipp = "BDCPP", 
                             dnbp = "DBuP", dphp = "DPhP")) |>
  dplyr::select(-c(unit, comment, barcode, barcode_creat, barcode_sg, 
            flg_match_creat, code_nr, flg_match_sg, analyte_lod, collection_month, creat)) |>
  mutate(flg_lod = if_else(is.na(flg_lod) | flg_lod != 1, 0, flg_lod),
         flg_nr = if_else(is.na(flg_nr) | flg_nr != 1, 0, flg_nr)) |>
  mutate(result = if_else(flg_lod == 1, 0.1/sqrt(2), result)) |>
  dplyr::select(-c(flg_nr, flg_lod)) |>
  #handling duplicate values for observations
  filter(participant_id %in% outcom_cov$participant_id) |>
  group_by(participant_id, visit, analyte_code) |>
  mutate(avg_result = mean(result)) |>
  distinct(participant_id, visit, analyte_code, .keep_all = TRUE) |>
  ungroup() |>
  dplyr::select(-result) |>
  rename(result = avg_result) 

#correlation between creatinine and urine speicific gravity : 0.66
#cor(ope_exp_raw$creat, ope_exp_raw$specific_gravity, use = "complete.obs")
```

#exposure data prep for aft imputation
```{r}
aft_imput_exp_prep<- ope_exp_raw  |>
  pivot_wider(names_from = analyte_code, values_from = result) |>
  mutate(across(3:7, ~ as.numeric(sprintf("%.3f", .)))) |>
  mutate(across(c(bcep, bdcipp, dnbp, dphp), ~ if_else(. < 0.1, 1, 0), .names = "ind_{.col}")) |>
  mutate(across(c(bcep, bdcipp, dnbp, dphp), ~ if_else(is.na(.), 1, 0), .names = "ind_na_{.col}")) |>
  mutate(across(8:11, ~ replace_na(., 0))) |>
  mutate(across(4:7, ~ replace_na(., 0.5)))  |>
  left_join(outcom_cov, by = "participant_id") |>
  mutate(across(where(is.character), as.factor))
```

#aft imputation for <LOD Values
```{r}
df_test<- aft_imput_exp_prep |>
  dplyr::select(c(4:11,28:30)) |>
  rename(
    z1 = bcep,
    z2 = bdcipp,
    z3 = dnbp,
    z4 = dphp,
    ind1 = ind_bcep,
    ind2 = ind_bdcipp,
    ind3 = ind_dnbp,
    ind4 = ind_dphp,
    x1 = rfr_gs_t,
    x2 = cotinine,
    x3 = lead
  ) |>
  drop_na() |>
  as.data.frame()

imputed_data <- perform_aft_imputation(
    data = df_test,                          # Your input dataset
    predictor_vars = c( "rfr_gs_t", "cotinine", "lead"),           # Predictor variables for the model
    z_vars = c("bcep", "bdcipp","dnbp", "dphp"),     # Z variables to be imputed
    ind_vars = c("ind_bcep", "ind_bdcipp","ind_dnbp", "ind_dphp"), # Indicator of <LOD values
    covariate_vars = c("bcep", "bdcipp","dnbp", "dphp")  # Covariates to be transformed
)
```


# get median specific gravity per visit
```{r}
median_sg_by_visit <- ope_exp_raw |>
  group_by(visit) |>
  summarise(median_sg = median(specific_gravity, na.rm = TRUE))
```

#count participants by visit and ope metabolite
```{r}
participant_counts <- ope_exp_raw |>
  group_by(visit, analyte_code) |>
  summarize(num_participants = n_distinct(participant_id)) |>
  pivot_wider(names_from = visit, values_from = num_participants)
```

#urine specific gravity standardization of OPE concentrations
```{r}
ope_sg_adj<- ope_exp_dist |>
  left_join(median_sg_by_visit, by = "visit") |>
  mutate(specific_gravity = if_else(is.na(specific_gravity), median_sg, specific_gravity))|>
  mutate(result_sg_adj= result * (median_sg - 1) / (specific_gravity - 1), na.rm=T,
         result_sg_adj = round(result_sg_adj, 3)) |>
  mutate(result_sg_adj = ifelse(flg_nr == 1, 9999, result_sg_adj))|> #observations with 9999 are NR due to interference
  mutate(result_sg_adj = ifelse(flg_lod == 1, 8888, result_sg_adj)) #observations with 8888 are below LOD

ope_sg_adj_wide<- ope_sg_adj |>
  select(c(participant_id, visit, analyte_code, result_sg_adj)) |>
  pivot_wider(id_cols = participant_id,
              names_from = c(analyte_code, visit), 
              values_from = result_sg_adj, 
              values_fill = 7777) #observations with 7777 indicate unmeasured
```

#exposure summary - pre-imputation
```{r}
exp_summary <- ope_sg_adj %>%
    group_by(analyte_code, visit) %>%
    summarize(
        median_iqr_pct_bel_lod_n = paste(
            round(quantile(result_sg_adj[result_sg_adj <= 500], 0.5, na.rm = TRUE), 2), "(", 
            round(quantile(result_sg_adj[result_sg_adj <= 500], 0.25, na.rm = TRUE), 2), "-", 
            round(quantile(result_sg_adj[result_sg_adj <= 500], 0.75, na.rm = TRUE), 2), ")",
            "[", 
            round((sum(flg_lod == 1) / n()) * 100, 2), "%;",
            n(), "]"
        )
    )

write_csv(exp_summary, paste0(res, "exp_summary_pre_imp.csv"))
```

#export exposure data
```{r}
write_csv(ope_sg_adj_wide, paste0(dat_loc, "exp_dat.csv"))
```



