---
title: "08_comparing_single_imp"
author: "Jagadeesh Puvvula"
date: "2025-01-08"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, haven, geepack, truncnorm, pracma, ggh4x, gtsummary)

dat_loc <- "~/Documents/ope_ssis/data/"
res<- "~/Documents/ope_ssis/result/"
```

#outcome and covariates data [n=236]
```{r}
outcom_cov<- read_csv("~/Documents/ope_ssis/res_sing_imput/out_cov.csv")
```

#ope data - pre and post natal exposure data received from aimin on Dec 13, 2023
```{r}
dat_1<- read_sas(paste0(dat_loc, "raw/dec_23_new_ver/ur_fr_sg_20210921.sas7bdat"))
dat_2<- read_sas(paste0(dat_loc, "raw/dec_23_new_ver/ur_fr_12y.sas7bdat"))

ope_dat<- bind_rows(dat_1, dat_2)

ope_exp_raw<- ope_dat |>
  clean_names() |>
  filter(!analyte_code %in% c("DBzP", "DCP", "TBBA", "BCPP", "iPPPP", "tBPPP"))|>
  group_by(visit) |>
  mutate(specific_gravity = if_else(is.na(specific_gravity) | specific_gravity == 1,  
                                    median(specific_gravity, na.rm = TRUE),  
                                    specific_gravity)) |>
  ungroup() |>
  mutate(visit = case_when(
    visit == "16W" ~ "prenatal_16W",
    visit == "26W" ~ "prenatal_26W",
    visit == "Birth" ~ "delivery",
    visit == "12M" ~ "postnatal_1_yr",
    visit == "24M" ~ "postnatal_2_yr",
    visit == "36M" ~ "postnatal_3_yr",
    visit == "60M" ~ "postnatal_5_yr",
    visit == "P3" ~ "postnatal_8_yr",
    visit == "P4" ~ "postnatal_12_yr",
    TRUE ~ as.character(visit)
  )) |>
  mutate(analyte_code = fct_recode(as.factor(analyte_code), 
                             bcep = "BCEtP", bdcipp = "BDCPP", 
                             dnbp = "DBuP", dphp = "DPhP")) |>
  select(-c(unit, comment, barcode, barcode_creat, barcode_sg, 
            flg_match_creat, code_nr, flg_match_sg, analyte_lod, collection_month, creat)) |>
  mutate(flg_lod = if_else(is.na(flg_lod) | flg_lod != 1, 0, flg_lod),
         flg_nr = if_else(is.na(flg_nr) | flg_nr != 1, 0, flg_nr)) |>
  #mutate(result = if_else(flg_lod == 1, 0.1/sqrt(2), result)) |>
  #select(-c(flg_nr, flg_lod)) |>
  #handling duplicate values for observations
  filter(participant_id %in% outcom_cov$participant_id) |>
  group_by(participant_id, visit, analyte_code) |>
  mutate(avg_result = mean(result)) |>
  distinct(participant_id, visit, analyte_code, .keep_all = TRUE) |>
  ungroup() |>
  select(-result) |>
  rename(result = avg_result) 



ope_exp_raw_si <- ope_exp_raw  |>
  pivot_wider(names_from = analyte_code, values_from = result) |>
  mutate(across(3:7, ~ as.numeric(sprintf("%.3f", .)))) |>
  mutate(across(c(bcep, bdcipp, dnbp, dphp), ~ if_else(. < 0.1, 1, 0), .names = "ind_{.col}")) |>
  mutate(across(c(bcep, bdcipp, dnbp, dphp), ~ if_else(is.na(.), 1, 0), .names = "ind_na_{.col}")) |>
  mutate(across(8:11, ~ replace_na(., 0))) |>
  mutate(across(4:7, ~ replace_na(., 0.5)))  |>
  left_join(outcom_cov, by = "participant_id") |>
  mutate(across(where(is.character), as.factor))
```

#exposure summary
```{r}
exp_summ <- ope_exp_raw %>%
  select(1:5) %>%
  distinct() %>%
  mutate(
    # Handle NAs first
    flg_nr = if_else(is.na(flg_nr), 0, flg_nr),
    flg_lod = if_else(is.na(flg_lod), 0, flg_lod),
    
    # Create mutually exclusive categories
    status = case_when(
      flg_nr == 1 ~ "Not Reported",
      flg_lod == 1 ~ "Below LOD",
      TRUE ~ "Detected"
    )
  ) %>%
  group_by(analyte_code, visit) %>%
  summarise(
    count = n(),
    not_reported = paste(sum(status == "Not Reported"), 
                         "(", round(mean(status == "Not Reported") * 100, 2), "%)", sep = ""),
    below_lod = paste(sum(status == "Below LOD"), 
                      "(", round(mean(status == "Below LOD") * 100, 2), "%)", sep = ""),
    detected = paste(sum(status == "Detected"), 
                     "(", round(mean(status == "Detected") * 100, 2), "%)", sep = ""),
    .groups = "drop"
  )
```

#detection pattern viz
```{r}
detect_viz <- ope_exp_raw |>
  select(1:5) |>
  mutate(participant_id = dense_rank(participant_id))|>
  distinct() |>
  mutate(
    # Handle NAs first
    flg_nr = if_else(is.na(flg_nr), 0, flg_nr),
    flg_lod = if_else(is.na(flg_lod), 0, flg_lod),
    
    # Create mutually exclusive categories
    status = case_when(
      flg_nr == 1 ~ "Not Reported",
      flg_lod == 1 ~ "Below LOD",
      TRUE ~ "Detected"
    )
  ) |>
  select(-c(4,5)) |>
  mutate(
  analyte_code = toupper(analyte_code),
  visit = recode(visit,
    "prenatal_16W" = "Gestational-16W",
    "prenatal_26W" = "Gestational-26W" ,
    "postnatal_1_yr" = "Child-1YR",
    "postnatal_2_yr" = "Child-2YR",
    "postnatal_3_yr" = "Child-3YR",
    "postnatal_5_yr" = "Child-5YR",
    "postnatal_8_yr" = "Child-8YR",
    "postnatal_12_yr" = "Child-12YR",
    .default = visit  # This keeps any unspecified levels as they are
  ),
  visit = str_to_title(visit)
  ) |>
  complete(
    analyte_code, 
    visit, 
    participant_id = unique(participant_id),  # Explicit 236 participants
    fill = list(status = "Unmeasured")
  ) |>
  mutate(analyte_visit = paste(analyte_code, visit, sep = "-"))

# Define the desired order for y-axis labels
custom_order <- c(
  "BCEP-Gestational-16w", "BCEP-Gestational-26w", "BCEP-Delivery",
  "BCEP-Child-1yr", "BCEP-Child-2yr", "BCEP-Child-3yr",
  "BCEP-Child-5yr", "BCEP-Child-8yr", "BCEP-Child-12yr",
  
  "BDCIPP-Gestational-16w", "BDCIPP-Gestational-26w", "BDCIPP-Delivery",
  "BDCIPP-Child-1yr", "BDCIPP-Child-2yr", "BDCIPP-Child-3yr",
  "BDCIPP-Child-5yr", "BDCIPP-Child-8yr", "BDCIPP-Child-12yr",
  
  "DNBP-Gestational-16w", "DNBP-Gestational-26w", "DNBP-Delivery",
  "DNBP-Child-1yr", "DNBP-Child-2yr", "DNBP-Child-3yr",
  "DNBP-Child-5yr", "DNBP-Child-8yr", "DNBP-Child-12yr",
  
  "DPHP-Gestational-16w", "DPHP-Gestational-26w", "DPHP-Delivery",
  "DPHP-Child-1yr", "DPHP-Child-2yr", "DPHP-Child-3yr",
  "DPHP-Child-5yr", "DPHP-Child-8yr", "DPHP-Child-12yr"
)

ggplot(detect_viz, aes(x = participant_id, 
                       y = factor(analyte_visit, levels = rev(custom_order)), 
                       fill = factor(status))) +
  geom_tile(width = 0.9, height = 0.9) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(limits = c(-1, 237), expand = c(0, 0)) +  # Specify x-axis limits
  geom_hline(yintercept = c(9.5, 18.5, 27.5, 36.5), 
             color = "grey10", linetype = "solid", size =1) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(values = c("Detected" = "#0072B2", "Below LOD" = "#E69F00", 
                               "Not Reported" = "#D55E00",
                               "Unmeasured" = "#999999")) +
  labs(title = "", x = "Study participants (n=236)", y = "Urinary OPE metabolite measurements") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_blank(),  
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.key.width = unit(0.8, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_blank(),
    plot.margin = margin(t = 0, r = 0, b = 0, l = 0)
  )

ggsave("~/Documents/ope_ssis/res_sing_imput/missing_pattern.tiff", 
       bg="white",
       width = 10,height = 6,
       dpi=300)
```

#Left-truncated single imputation
```{r}
#impute with LOD/sqrt-2
ope_exp_raw <- ope_exp_raw |> 
  mutate(result = ifelse(flg_lod == 1, 0.1/sqrt(2), result),
         log2_result = log2(result))

#generate distribution summary
LODmeansd.all <- ope_exp_raw |> 
  group_by(analyte_code, visit) |> 
  summarise(meanlog2 = mean(log2_result, na.rm= TRUE), sdlog2 = sd(log2_result, na.rm=TRUE)) |> 
  mutate(test.name = (paste0(analyte_code, ".t", visit)))

# Create all combinations of analyte_code and visit
combinations <- expand.grid(
  analyte_code = unique(ope_exp_raw$analyte_code),
  visit = unique(ope_exp_raw$visit)
)

#perform imputation
results <- combinations |>
  pmap_dfr(~ process_combination(.x, .y)) |>
  rename(imputed_res = analyte_result) |>
  select(c(1:4))
```

#add single imputations to ope_data
#re-load the ope_exp_raw data to join
```{r}
ope_df<- results |>
  right_join(ope_exp_raw, by = c("participant_id", "analyte_code", "visit")) |>
  do({
    # For each visit group, calculate the median of specific_gravity
    median_sg <- median(.$specific_gravity, na.rm = TRUE)
    
    # Apply the mutate operation for result_sg_adj and round it
    . |> 
      mutate(
        result_sg_adj = imputed_res * (median_sg - 1) / (specific_gravity - 1),
        result_sg_adj = round(result_sg_adj, 3)
      )
  }) |> 
  ungroup() 
```

#export clean exposure data
```{r}
write_csv(ope_df, "~/Documents/ope_ssis/res_sing_imput/ope_processed_df_feb182025.csv")
```

#participant volume by visit and ICC calculation
```{r}
ope_df<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_processed_df_feb182025.csv") |>
  select(c(1:3,10)) 

#full summary
summary_table <- ope_df |>
    mutate(period = if_else(str_starts(visit, "postnatal"),
                            "child",
                            "gestation")) |>
    distinct(participant_id, visit, period) |>
    group_by(participant_id, period) |>
    summarise(n_visits = n(), .groups = "drop") |>
    pivot_wider(names_from = period, values_from = n_visits, values_fill = 0) |>
    mutate(lifetime = gestation + child) |>
    count(gestation, child, lifetime, name = "n_participants")

#for publication summary
summary_table <- ope_df |>
  mutate(period = if_else(str_starts(visit, "postnatal"),
                          "child",
                          "gestation")) |>
  distinct(participant_id, visit, period) |>
  group_by(participant_id, period) |>
  summarise(n_visits = n(), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = n_visits, values_fill = 0) |>
  mutate(
    lifetime = gestation + child,
    gestation_group = if_else(gestation < 2, "<2", "≥2"),
    child_group = if_else(child < 4, "<4", "≥4"),
    lifetime_group = if_else(lifetime < 6, "<6", "≥6")
  ) |>
  count(gestation_group, child_group, lifetime_group, name = "n_participants")


#ICC
ope_icc<- ope_df |>
  mutate(
    visit = recode(visit,
                   "prenatal_16W" = "v1", "prenatal_26W" = "v2", "delivery" = "v3",
                   "postnatal_1_yr" = "v4", "postnatal_2_yr" = "v5", "postnatal_3_yr" = "v6",
                   "postnatal_5_yr" = "v7", "postnatal_8_yr" = "v8",  "postnatal_12_yr" = "v9"  )) |>
  filter(!(analyte_code == "dnbp" & visit %in% c("v4", "v5"))) |>
  filter(visit %in% c("v1", "v2", "v3", "v4", "v5", "v6", "v7", "v8", "v9")) 

#"v1", "v2", "v3",


library(lme4)

calc_icc_lme <- function(df, analyte) {
  
  # Auto-detect visits present in the data for this analyte
  visits_present <- df %>%
    filter(analyte_code == analyte, 
           result_sg_adj > 0,
           !is.na(result_sg_adj)) %>%
    distinct(visit) %>%
    pull(visit) %>%
    sort()
  
  cat("\n--- Analyte:", analyte, "---\n")
  cat("Visits detected:", paste(visits_present, collapse = ", "), "\n")
  cat("N visits:", length(visits_present), "\n")
  
  dat <- df %>%
    filter(
      analyte_code == analyte,
      visit %in% visits_present,
      result_sg_adj > 0,
      !is.na(result_sg_adj)
    ) %>%
    mutate(log_result = log(result_sg_adj))
  
  cat("N rows:", nrow(dat), "\n")
  cat("Var (raw):", round(var(dat$result_sg_adj), 3), "\n")
  cat("Var (log):", round(var(dat$log_result), 3), "\n")
  
  model <- lmer(log_result ~ 1 + (1 | participant_id),
                data = dat,
                REML = TRUE,
                control = lmerControl(
                  optimizer = "bobyqa",
                  optCtrl   = list(maxfun = 2e5)
                ))
  
  vc          <- as.data.frame(VarCorr(model))
  var_between <- vc$vcov[1]
  var_within  <- vc$vcov[2]
  icc         <- var_between / (var_between + var_within)
  
  tibble(
    analyte_code    = analyte,
    visits_used     = paste(visits_present, collapse = ", "),  # track which visits were used
    n_visits        = length(visits_present),
    n_obs           = nrow(dat),
    n_participants  = n_distinct(dat$participant_id),
    var_between     = round(var_between, 3),
    var_within      = round(var_within, 3),
    ICC             = round(icc, 2),
    reproducibility = case_when(
      icc <= 0.40              ~ "Poor",
      icc > 0.40 & icc < 0.75 ~ "Fair to Good",
      icc >= 0.75              ~ "Excellent"
    )
  )
}

# --- Run --- 
analytes        <- unique(ope_icc$analyte_code)
icc_lme_results <- map_dfr(analytes, ~ calc_icc_lme(ope_icc, .x))
print(icc_lme_results)
```

#visualize concentrations by visit
```{r}
ope_viz <- ope_df |>
  mutate(
    analyte_code = toupper(analyte_code),
    visit = recode(visit, "prenatal_16W" = "16W", "prenatal_26W" = "26W", "delivery" = "Delivery",
                   "postnatal_1_yr" = "1Y", "postnatal_2_yr" = "2Y", "postnatal_3_yr" = "3Y",
                   "postnatal_5_yr" = "5Y", "postnatal_8_yr" = "8Y",  "postnatal_12_yr" = "12Y"),
    visit = factor(visit, levels = c("16W","26W","Delivery","1Y","2Y","3Y","5Y","8Y","12Y")),
    stage = case_when(
      visit %in% c("16W","26W","Delivery") ~ "Gestation",
      TRUE ~ "Childhood"
    )
  ) |>
  filter(!(analyte_code == "DNBP" & visit %in% c("1Y", "2Y")))

ggplot(ope_viz, aes(x = visit, y = log2(result_sg_adj),
                    group = participant_id, color = stage)) +
  geom_line(alpha = 0.1) +
  geom_point(alpha = 0.1, size = 1) +
  facet_wrap(~ analyte_code, scales = "free_y") +
  scale_color_manual(
    values = c("Gestation" = "#D55E00",
               "Childhood" = "#0072B2"),
    guide = guide_legend(override.aes = list(alpha = 1, size = 3))
  ) +
  theme_minimal() +
  labs(x = "Visit",
       y = "OPE Concentration-Log2 & SG-standardized",
       color = "Exposure Window") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom"
  )

ggsave("~/Documents/ope_ssis/exp_by_visit.tiff", 
       width = 7, height = 6, 
       dpi = 300, bg = "white")
```


#join exposures with outcome and covariates
```{r}
ope_df<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_processed_df_feb182025.csv") |>
  select(c(1:3,10)) |>
  pivot_wider(
    names_from = analyte_code,
    values_from = result_sg_adj
  )

analy_df<- right_join(ope_df, outcom_cov, by="participant_id")

write_csv(analy_df, "~/Documents/ope_ssis/res_sing_imput/ope_ssis_fin_feb182025.csv")
```

#visualizing repeated exposures
```{r}
#visualize data
viz_df <- ope_ssis |> 
  select(c(1:6)) |>
  mutate(visit_numeric = as.numeric(gsub("V", "", visit))) |> 
  pivot_longer(cols = c(bcep, bdcipp, dnbp, dphp), names_to = "chemical", values_to = "value") |>
  filter(!(chemical == "dnbp" & visit %in% c("V4", "V5"))) |>
  filter(value < 250) |>
  mutate(chemical = toupper(chemical))  # Convert 'chemical' to uppercase

ggplot(viz_df, aes(x = visit_numeric, 
                   y = value, 
                   color = chemical, 
                   group = interaction(participant_id, chemical))) +
  geom_line(alpha = 0.5, linewidth = 0.25) +
  geom_point(alpha = 0.25) +
  facet_grid(chemical~., scale = "free")+
  stat_summary(aes(group = chemical), fun = median, geom = "line", 
               linewidth = 1., linetype = "solid", color = "grey25") +
  scale_y_continuous(trans = "log2", labels = label_number(accuracy = 0.01)) +
  scale_x_continuous(breaks = 1:9, labels = paste0("V", 1:9)) +
  labs(x = "Visit", y = "Concentration", color = "OPE biomarker") +
  theme_minimal()+
  theme(legend.position = "bottom")

ggsave("~/Documents/ope_ssis/res_sing_imput/exp_ovr_time.tiff",
       width = 9,
       height = 6,
       bg = "white",
       dpi=300)
```

# Complete solution with proper handling of all edge cases
# Calculate AUC for multiple variables (bcep, bdcipp, dnbp, dphp)
# Use calculation suggested by aimin: https://trumpexcel.com/calculate-area-under-curve-excel/ 
```{r}
#gestational age | Some participants dont have NA which is again creating a problem
ga_w<- haven::read_sas("~/Documents/prenatal_OPFR/outcome_ga.sas7bdat") |>
  clean_names() |>
  drop_na() |>
  mutate(ga = as.integer(ga))

#summary of ope exposures for supplement table
read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_fin_feb182025.csv") |>
  mutate(across(where(is.character), as.factor)) |>
  mutate(dnbp = dplyr::case_when(visit %in% c("postnatal_1_yr", "postnatal_2_yr") ~ NA_real_, 
                                TRUE ~ dnbp)) |>
  select(c(2:6)) |>
  tbl_summary(
    by=visit,
    
  )

# Read and prepare data
ope_ssis <- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_fin_feb182025.csv") |>
  mutate(across(where(is.character), as.factor)) |>
  mutate(dnbp = dplyr::case_when(visit %in% c("postnatal_1_yr", "postnatal_2_yr") ~ NA_real_, 
                                TRUE ~ dnbp)) |>
  mutate(
    visit = recode(visit, 
                  "prenatal_16W" = "16",
                  "prenatal_26W" = "26"),
    
    visit = case_when(
    visit == "delivery" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)])
    ),
    visit == "postnatal_1_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 52
    ),
    visit == "postnatal_2_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 104
    ),
    visit == "postnatal_3_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 156
    ),
    visit == "postnatal_5_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 260
    ),
    visit == "postnatal_8_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 416
    ),
    visit == "postnatal_12_yr" ~ as.character(
      ifelse(is.na(ga_w$ga[match(participant_id, ga_w$participant_id)]),
             39,
             ga_w$ga[match(participant_id, ga_w$participant_id)]) + 624
    ),
    TRUE ~ visit
  ),
  visit = as.numeric(visit)
  )|>
  select(participant_id, visit, bcep, bdcipp, dnbp, dphp)|>
  filter(visit>80)

# Calculate AUC - applying logic as discussed with aimin
calculate_auc <- function(times, concentrations) {
  # Remove NA concentrations and corresponding times
  valid_data <- data.frame(time = times, conc = concentrations) %>%
    filter(!is.na(conc))
  
  # Sort data by time
  ordered_data <- valid_data %>% arrange(time)
  
  # If no valid points, return NA
  if(nrow(ordered_data) == 0) {
    return(NA_real_)
  }
  
  # If only one valid point, return that concentration value directly
  if(nrow(ordered_data) == 1) {
    return(ordered_data$conc[1])
  }
  
  # Calculate AUC using trapezoidal rule
  auc <- 0
  for(i in 1:(nrow(ordered_data)-1)) {
    time_diff <- ordered_data$time[i+1] - ordered_data$time[i]
    avg_conc <- (ordered_data$conc[i+1] + ordered_data$conc[i]) / 2
    auc <- auc + (time_diff * avg_conc)
  }
  
  # Normalize AUC by total time span (last visit - first visit)
  time_span <- max(ordered_data$time) - min(ordered_data$time)
  normalized_auc <- auc / time_span
  
  return(normalized_auc)
}

# Apply the AUC calculation to your dataset directly
calculate_metabolite_aucs <- function(data) {
  # Get unique participant IDs
  participants <- unique(data$participant_id)
  
  # Initialize results dataframe
  results <- data.frame(
    participant_id = numeric(),
    bcep_auc = numeric(),
    bdcipp_auc = numeric(),
    dnbp_auc = numeric(),
    dphp_auc = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Calculate AUC for each participant and each metabolite
  for(pid in participants) {
    participant_data <- data %>% filter(participant_id == pid)
    
    bcep_auc <- calculate_auc(participant_data$visit, participant_data$bcep)
    bdcipp_auc <- calculate_auc(participant_data$visit, participant_data$bdcipp)
    dnbp_auc <- calculate_auc(participant_data$visit, participant_data$dnbp)
    dphp_auc <- calculate_auc(participant_data$visit, participant_data$dphp)
    
    results <- rbind(results, data.frame(
      participant_id = pid,
      bcep_auc = bcep_auc,
      bdcipp_auc = bdcipp_auc,
      dnbp_auc = dnbp_auc,
      dphp_auc = dphp_auc
    ))
  }
  
  return(results)
}

# Run the function on your dataset
auc_results <- calculate_metabolite_aucs(ope_ssis)
```

#OPE estimation by trapezoid method
```{r}
#life-time OPE exposure - trapezoid estimation
ope_life<- auc_results |> rename_with(~ paste0(.x, "_life"), ends_with("auc"))

#gestational OPE exposure - trapezoid estimation
ope_gest<- auc_results |> rename_with(~ paste0(.x, "_gest"), ends_with("auc"))

#gestational OPE exposure - trapezoid estimation
ope_child<- auc_results |> rename_with(~ paste0(.x, "_child"), ends_with("auc"))

#join different sections
ope_exposures <- reduce(list(ope_life, ope_gest, ope_child), full_join, by = "participant_id")

#created on March 21 - calculations as recommended by Aimin
write_csv(ope_exposures, "~/Documents/ope_ssis/res_sing_imput/ope_trap_aggr.csv")
```

#outcome correlation heatmap
```{r}
cov_out<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_fin_feb182025.csv") |>
  select(c(1,7:11)) |>
  distinct() |>
  rename(`Self-reported social skill` = "ss_std_score_c",
         `Caregiver-reported social skill` = "ss_std_score_p",
         `Self-reported problem behavior` = "pb_std_score_c",
         `Caregiver-reported problem behavior` = "pb_std_score_p")

# Compute correlations for Overall, Male, and Female
cor_overall_df <- cov_out |>
  select(all_of(vars)) |>
  cor(use = "pairwise.complete.obs", method = "spearman") |>
  melt() |>
  mutate(group = "Overall") |>
  rowwise() |>
  mutate(p_value = cor.test(cov_out[[Var1]], cov_out[[Var2]], method = "spearman", exact = FALSE)$p.value)

cor_male_df <- cov_out |>
  filter(gender == "Male") |>
  select(all_of(vars)) |>
  cor(use = "pairwise.complete.obs", method = "spearman") |>
  melt() |>
  mutate(group = "Male") |>
  rowwise() |>
  mutate(p_value = cor.test(cov_out[[Var1]], cov_out[[Var2]], method = "spearman", exact = FALSE)$p.value)

cor_female_df <- cov_out |>
  filter(gender == "Female") |>
  select(all_of(vars)) |>
  cor(use = "pairwise.complete.obs", method = "spearman") |>
  melt() |>
  mutate(group = "Female") |>
  rowwise() |>
  mutate(p_value = cor.test(cov_out[[Var1]], cov_out[[Var2]], method = "spearman", exact = FALSE)$p.value)

# Combine into single dataframe
cor_all <- bind_rows(cor_overall_df, cor_male_df, cor_female_df)

# Ensure proper factor levels to control the order of the facets
cor_all$group <- factor(cor_all$group, levels = c("Overall", "Male", "Female"))

# Add rounded coefficients and significance based on p-value
cor_all <- cor_all %>%
  mutate(
    coefficient = round(value, 2),
    significance = ifelse(p_value < 0.05, "*", "")
  )

# Plot
ggplot(cor_all, aes(Var1, Var2, fill = coefficient)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = "#DC143C", mid = "white", high = "#0047AB",
    midpoint = 0, limit = c(-1, 1), space = "Lab",
    name = "Spearman ρ"
  ) +
  facet_wrap(~ group, nrow = 1) +  # Separate by 'group' (Overall, Male, Female)
  coord_fixed() +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 20, hjust = 1),
    panel.spacing = unit(0.5, "lines"),
    legend.position = "bottom",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.2, "cm")
  ) +
  # Add text for coefficients and significance
  geom_text(aes(label = paste(coefficient, significance)), color = "black", size = 4)

ggsave("~/Documents/ope_ssis/res_sing_imput/ssis_corr.tiff", 
       width = 12,height = 5,
       dpi=300,
       bg="white")

#trends
# Reshape data to long format
cov_long <- cov_out |>
  pivot_longer(
    cols = c(`Self-reported social skill`, `Caregiver-reported social skill`,
             `Self-reported problem behavior`, `Caregiver-reported problem behavior`),
    names_to = "measure_type",
    values_to = "score"
  ) |>
  mutate(
    domain = case_when(
      str_detect(measure_type, "social skill") ~ "Social Skill",
      str_detect(measure_type, "problem behavior") ~ "Problem Behavior"
    ),
    reporter = case_when(
      str_detect(measure_type, "Self") ~ "Self",
      str_detect(measure_type, "Caregiver") ~ "Caregiver"
    ),
    reporter = factor(reporter, levels = c("Self", "Caregiver"))
  )

cov_long2 <- cov_long |> 
  mutate(gender_facet = gender) |> 
  bind_rows(
    cov_long |> 
      mutate(gender_facet = "Overall")
  ) |>
  mutate(gender_facet = factor(gender_facet, levels = c("Overall", "Male", "Female")))

ggplot(cov_long2, aes(x = reporter, y = score, group = participant_id)) +
  geom_line(aes(color = gender), alpha = 0.2, linewidth = 0.8) +
  #stat_summary(aes(group = gender, color = gender), fun = median, geom = "line", linewidth = 1.5) +
  facet_grid(domain ~ gender_facet, scales = "free_y") +
  theme_minimal(base_size = 8) +
  labs(
    x = "Reporter",
    y = "Score",
    color = "Adolescent sex",
    title = " "
  ) +
  theme(
    strip.text = element_text(size = 6, face = "bold"),
    legend.position = "bottom"
  )

ggsave("~/Documents/ope_ssis/res_sing_imput/ssis_trends.tiff", 
       width = 7,height = 4,
       dpi=300,
       bg="white")
```



#qgcomp with aggregated data
```{r}
ope<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_trap_aggr.csv")

cov_out<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_fin_feb182025.csv") |>
  mutate(across(where(is.character), as.factor)) |>
  mutate(#across(3:6, log2),
         visit = recode(visit, 
                        "prenatal_16W" = "V1",
                        "prenatal_26W" = "V2",
                        "delivery" = "V3",
                        "postnatal_1_yr" = "V4",
                        "postnatal_2_yr" = "V5",
                        "postnatal_3_yr" = "V6",
                        "postnatal_5_yr" = "V7",
                        "postnatal_8_yr" = "V8",
                        "postnatal_12_yr" = "V9"),
         visit = factor(visit, levels = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9"))
         ) |>
  select(-c(2:6)) |>
  distinct()

analy_df<- left_join(ope, cov_out, by="participant_id")

write_csv(analy_df, "~/Documents/ope_ssis/res_sing_imput/ope_ssis_mar2025.csv")
```

#qgcomp
```{r}
analy_df<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_mar2025.csv") |>
  mutate(cotinine = log2(cotinine),
         lead = log2(lead),
         cohort = "HOME") |>
  mutate_if(is.character, as.factor) |>
  rename(sex= gender)

#AUC summary - Table 2
df_summary <- analy_df |>
    select(2:13) |>
    pivot_longer(everything(), names_to = "variable", values_to = "value") |>
    group_by(variable) |>
    summarise(
        median_iqr = sprintf(
            "%.2f (%.2f, %.2f)",
            median(value, na.rm = TRUE),
            quantile(value, 0.25, na.rm = TRUE),
            quantile(value, 0.75, na.rm = TRUE)
        ),
        mean_sd = sprintf(
            "%.2f (%.2f)",
            mean(value, na.rm = TRUE),
            sd(value, na.rm = TRUE)
        ),
        geom_mean_sd = if (all(value > 0, na.rm = TRUE)) {
            sprintf(
                "%.2f (%.2f)",
                exp(mean(log(value), na.rm = TRUE)),
                exp(sd(log(value), na.rm = TRUE))
            )
        } else {
            NA_character_
        },
        .groups = "drop"
    )

  
#mixture<- names(analy_df)[6:9] #gesta (n=236)

#for childhood exposure analysis (n=232)
analy_df_chld <- analy_df |> select(c(1,10:29)) |>drop_na()
mixture<- names(analy_df_chld)[2:5] #childhood


qgcomp_func(outcomes = c("ss_std_score_c", "pb_std_score_c", "ss_std_score_p", "pb_std_score_p"), 
            data = analy_df_chld, #update this step
            output_folder = "~/Documents/ope_ssis/res_sing_imput/qgc/chld_auc", 
            include_sex = TRUE, 
            chemicals = list("bcep_auc_child",   "bdcipp_auc_child", "dnbp_auc_child",   "dphp_auc_child"  ),
            covariates = list("child_race",
                              "mat_age", "mari_st_p4", "mom_edu_p4_cat",
                              "mid_income_p4","bdi_p4", "depression_p4_cat", "rfr_gs_t",
                              "cotinine", "lead"),
            q = 4, b=400, seed = 2025)
```

#adolescent sex interaction
```{r}
mixture<- names(analy_df_chld)[2:5] #gestational

qgcomp_pval_extract <- function(outcomes, data, mixture, q) {
  results <- lapply(outcomes, function(outcome) {
    
    #bcep_auc_gest + bdcipp_auc_gest + dnbp_auc_gest + dphp_auc_gest + 
    #bcep_auc_life + bdcipp_auc_life + dnbp_auc_life + dphp_auc_life +
    
    formula <- as.formula(paste(outcome, "~ 
    bcep_auc_child + bdcipp_auc_child + dnbp_auc_child + dphp_auc_child +
    child_race + mat_age + mari_st_p4 + mom_edu_p4_cat + mid_income_p4 + bdi_p4 +
    depression_p4_cat + rfr_gs_t + cotinine + lead"))
    
    result <- qgcomp.emm.noboot(formula, 
                                expnms = mixture, 
                                emmvar = "sex", 
                                data = data, 
                                family = gaussian(), 
                                q = q)$pval
    
    # Return the last p-value
    last_pval <- result[length(result)]
    
    return(list(
      outcome = outcome,
      pval = last_pval
    ))
  })
  
  # Print results
  for(res in results) {
    cat("Outcome:", res$outcome, "| P-value:", res$pval, "\n")
  }
  
  # Return results invisibly
  return(results)
}

# Example usage
mixture <- names(analy_df_chld)[2:5] #gestational
qgcomp_pval_extract(
  outcomes = c("ss_std_score_c", "pb_std_score_c", "ss_std_score_p", "pb_std_score_p"), 
  data = analy_df_chld,
  mixture = mixture,
  q = 4
)

#use qgcompint version 0.7.0 and convert dataset to a dataframe using as.data.frame
qgcompint::qgcomp.emm.boot(
  f = ss_std_score_c ~ sex + bcep_auc_life + bdcipp_auc_life + dnbp_auc_life + dphp_auc_life + 
    child_race + mat_age + mari_st_p4 + mom_edu_p4_cat + mid_income_p4 + 
    bdi_p4 + depression_p4_cat + rfr_gs_t + cotinine + lead, 
  expnms = c('bcep_auc_life', 'bdcipp_auc_life', 'dnbp_auc_life', 'dphp_auc_life'), 
  emmvar = "sex", 
  data = analy_df, 
  family = gaussian(), 
  B=400, 
  seed = 2025,
  q = 4)
```


#get estimates
```{r}
results_child<- get_gcomp_estimates("~/Documents/ope_ssis/res_sing_imput/qgc/chld_auc")|> 
  mutate(auc_category = "Childhood") |> select(-c(1,2))
results_gesta<- get_gcomp_estimates("~/Documents/ope_ssis/res_sing_imput/qgc/gesta_auc")|> 
  mutate(auc_category = "Gestational") |> select(-c(1,2))
results_life<- get_gcomp_estimates("~/Documents/ope_ssis/res_sing_imput/qgc/life_auc")|> 
  mutate(auc_category = "Lifetime")|> select(-c(1,2))

est_comb<- bind_rows(results_child, results_gesta, results_life) |>
  mutate(
    gender = case_when(
    str_starts(filename, "nb_Male") ~ "Male",
    str_starts(filename, "boot_Male") ~ "Male",
    str_starts(filename, "nb_Female") ~ "Female",
    str_starts(filename, "boot_Female") ~ "Female",
    str_starts(filename, "nb_all") ~ "All",
    str_starts(filename, "boot_all") ~ "All"
  ),
  group = case_when(
      str_ends(filename, "_c.rda") ~ "adolescent",
      str_ends(filename, "_p.rda") ~ "caregiver",
      TRUE ~ "Unknown group"
    ),
  boot_strp = case_when(
      str_starts(filename, "nb_") ~ "no_boot",
      str_starts(filename, "boot_") ~ "boot",
      TRUE ~ "Unknown group"
    ),
    outcome = case_when(
      str_detect(filename, "pb_std_score") ~ "problem behavior",
      str_detect(filename, "ss_std_score") ~ "social skill",
      TRUE ~ "Unknown behavior"
    ),
    ci_null = if_else(p_value > 0.05, "not significant", "significant")
  )
```

#Export estimates
```{r}
write_csv(est_comb, "~/Documents/ope_ssis/res_sing_imput/qgc/qgc_estimates_all_auc_categ.csv")
```

#plot qgc results
```{r}
res_plt<- read_csv("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_estimates_all_auc_categ.csv") |>
  mutate(
    across(c(group, outcome, auc_category, gender), str_to_title),
    auc_category = fct_relevel(auc_category, "Gestational", "Childhood", "Lifetime"))

cbbPalette <- c("#D55E00", "#0072B2")

ggplot(res_plt, aes(x = as.factor(group), y = estimate, ymin = lower_ci, ymax = upper_ci)) + 
  geom_pointrange(aes(col = as.factor(ci_null), shape=boot_strp), 
                  position=position_dodge(width=0.5),size = 0.6) +
  geom_hline(aes(yintercept = 0), linetype="dashed") + 
  scale_colour_manual(values=cbbPalette) +
  labs(x=NULL,
       y=expression(paste(beta[Q-gcomp]," [95% CI]")),
       title = "",
       caption = "")+
  theme_bw()+
  theme(axis.line = element_line(colour = "black"))+
  theme(plot.caption = element_text(size = 6),
        plot.title=element_text(size = 10, hjust = 0),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        text=element_text(size=10),
        axis.text = element_text(size = 10),
        axis.text.x=element_text(size=10,angle=0),
        axis.ticks.x=element_blank(),
        strip.text.x=element_text(size=10,angle=0),
        strip.text.y = element_text(size=10),
        strip.placement = "outside",
        strip.background = element_rect(fill=NULL, colour = NULL),
        legend.box.just = "center",
        legend.spacing = unit(0, "cm"),
        legend.position = "bottom",
        legend.box = "horizontal",
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.key = element_blank(),
        legend.key.height = unit(3, "mm"))+
  guides(col="none",
         shape=guide_legend(title = "Bootstrap"))+
  ggh4x::facet_nested(gender~outcome+auc_category, scales = "free", switch = "x", space = "free_x")
```

```{r}
ggsave("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_auc_all.tiff", 
       width = 11,height = 7,
       dpi=300 )
```

#qgcomp weights
```{r}
qgc_wt_chld <-  extract_weights("~/Documents/ope_ssis/res_sing_imput/qgc/chld_auc") |> 
  mutate(auc_category = "Childhood")
qgc_wt_gesta <-  extract_weights("~/Documents/ope_ssis/res_sing_imput/qgc/gesta_auc") |> 
  mutate(auc_category = "Gestational")
qgc_wt_life <-  extract_weights("~/Documents/ope_ssis/res_sing_imput/qgc/life_auc") |> 
  mutate(auc_category = "Lifetime")

qgc_wt_comb<- bind_rows(qgc_wt_chld, qgc_wt_gesta, qgc_wt_life) |>
  mutate(
    gender = case_when(
    str_starts(filename, "nb_Male") ~ "Male",
    str_starts(filename, "boot_Male") ~ "Male",
    str_starts(filename, "nb_Female") ~ "Female",
    str_starts(filename, "boot_Female") ~ "Female",
    str_starts(filename, "nb_all") ~ "All",
    str_starts(filename, "boot_all") ~ "All"
  ),
  group = case_when(
      str_ends(filename, "_c") ~ "adolescent",
      str_ends(filename, "_p") ~ "caregiver",
      TRUE ~ "Unknown group"
    ),
    outcome = case_when(
      str_detect(filename, "pb_std_score") ~ "problem behavior",
      str_detect(filename, "ss_std_score") ~ "social skill",
      TRUE ~ "Unknown behavior"
    ),
    chemical = str_extract(chemical, "^[^_]+") %>%
           str_to_upper(),
    weight= if_else(direction=="neg", weight*-1, weight)
  ) |>
  mutate_if(is.numeric, ~ as.numeric(sprintf("%.2f", .)))

#save weights
rm(qgc_wt_chld, qgc_wt_gesta, qgc_wt_life)

write_csv(qgc_wt_comb, "~/Documents/ope_ssis/res_sing_imput/qgc/qgc_all_weigh.csv")
```

#QGC estimates and weight table for manuscript
```{r}
res_plt<- read_csv("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_estimates_all_auc_categ.csv")|>
  mutate(
    filename = str_remove(filename, "\\.rda$"),
    across(c(group, outcome, auc_category, gender), str_to_title),
    result = sprintf("%.2f (%.2f, %.2f)", estimate, lower_ci, upper_ci),
    filename = str_replace(filename, "^boot_|^nb_", ""),
    filename= paste(filename, auc_category, sep = "_")
         ) |> 
  select(c(1,9,12)) |>
  pivot_wider(names_from = "boot_strp", values_from = "result")

#weights
wt<- qgc_wt_comb |>
  mutate(
    filename= paste(filename, auc_category, sep = "_"),
    filename = str_replace(filename, "^boot_|^nb_", "")
  )|> 
  select(c(4,1,2))|> 
  pivot_wider(names_from = "chemical", values_from = "weight") |>
  select(c(1,8,7,6,3,4,5,2))

#reorder
custom_order <- c(
  "all_proBEH_adole_Gestational", "all_proBEH_adole_Childhood",  "all_proBEH_adole_Lifetime",
  "Female_proBEH_adole_Gestational",  "Female_proBEH_adole_Childhood", "Female_proBEH_adole_Lifetime",
  "Male_proBEH_adole_Gestational", "Male_proBEH_adole_Childhood", "Male_proBEH_adole_Lifetime",
  
  "all_proBEH_careg_Gestational", "all_proBEH_careg_Childhood",  "all_proBEH_careg_Lifetime",
  "Female_proBEH_careg_Gestational",  "Female_proBEH_careg_Childhood", "Female_proBEH_careg_Lifetime",
  "Male_proBEH_careg_Gestational", "Male_proBEH_careg_Childhood", "Male_proBEH_careg_Lifetime",
  
  "all_socSKI_adole_Gestational", "all_socSKI_adole_Childhood",  "all_socSKI_adole_Lifetime",
  "Female_socSKI_adole_Gestational",  "Female_socSKI_adole_Childhood", "Female_socSKI_adole_Lifetime",
  "Male_socSKI_adole_Gestational", "Male_socSKI_adole_Childhood", "Male_socSKI_adole_Lifetime",
  
  "all_socSKI_careg_Gestational", "all_socSKI_careg_Childhood",  "all_socSKI_careg_Lifetime",
  "Female_socSKI_careg_Gestational",  "Female_socSKI_careg_Childhood", "Female_socSKI_careg_Lifetime",
  "Male_socSKI_careg_Gestational", "Male_socSKI_careg_Childhood", "Male_socSKI_careg_Lifetime"
)
  
#result table formatted
tbl<- left_join(res_plt, wt, by = "filename") |>
  select(-c(4)) |>
  mutate(filename = case_when(
    # Replace longest patterns first
    str_detect(filename, "_pb_std_score_") ~ str_replace(filename, "_pb_std_score_", "_proBEH_"),
    str_detect(filename, "_ss_std_score_") ~ str_replace(filename, "_ss_std_score_", "_socSKI_"),
    TRUE ~ filename
  )) |>
  mutate(filename = case_when(
    str_detect(filename, "_c_") ~ str_replace(filename, "_c_", "_adole_"),
    str_detect(filename, "_p_") ~ str_replace(filename, "_p_", "_careg_"),
    TRUE ~ filename
  )) |>
  mutate(filename = factor(filename, levels = custom_order)) |>
  arrange(filename)

write_csv(tbl, "~/Documents/ope_ssis/res_sing_imput/qgc/qgc_mansc_tbl.csv")
```

#OPE - ssis linear regression
```{r}
analy_df<- read_csv("~/Documents/ope_ssis/res_sing_imput/ope_ssis_mar2025.csv") |>
  mutate(cotinine = log2(cotinine),
         lead = log2(lead),
         cohort = "HOME") |>
  mutate_if(is.character, as.factor) |>
  rename(sex= gender) |>
   mutate(across(2:13, log2))

lm_df<- lm_func(
  dependent_vars = c("ss_std_score_c",    "pb_std_score_c",
                     "ss_std_score_p",    "pb_std_score_p"),
  independent_vars=c("bcep_auc_life", "bdcipp_auc_life", "dnbp_auc_life", "dphp_auc_life", 
                     "bcep_auc_gest", "bdcipp_auc_gest", "dnbp_auc_gest", "dphp_auc_gest",
                     "bcep_auc_child", "bdcipp_auc_child",  "dnbp_auc_child", "dphp_auc_child"),
  covariates = c("child_race",  "mat_age", "mari_st_p4", "mom_edu_p4_cat", "mid_income_p4",
                 "bdi_p4", "depression_p4_cat", "rfr_gs_t", "cotinine", "lead"),
  dat = analy_df, 
  include_sex = TRUE
  )
```

```{r}
lm_res_tbl <- lm_df %>%
  mutate(result = sprintf("%.2f (%.2f, %.2f)", coefficient, ci_lower, ci_upper)) %>%
  # Make sure all combinations exist
  complete(independent_variable, dependent_variable, sex_level, fill = list(result = NA)) %>%
  pivot_wider(id_cols = independent_variable,
              names_from = c(dependent_variable, sex_level),
              values_from = result)

write_csv(lm_res_tbl, "~/Documents/ope_ssis/res_sing_imput/mlr_results.csv")
```


#main figure
```{r}
res_plt<- read_csv("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_estimates_all_auc_categ.csv") |>
  mutate(
    across(c(group, outcome, auc_category, gender), str_to_title),
    auc_category = fct_relevel(auc_category, "Gestational", "Childhood", "Lifetime")) |>
   filter(str_starts(filename, "boot_"))

qgc_wt<- read_csv("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_mansc_tbl.csv") |>
  select(-c(3:5)) |>
  separate(filename, into = c("group", "outcome", "reporter", "exposure_period"), sep = "_") %>%
  mutate(
    group = case_when(
      group == "all" ~ "Overall",
      group == "Male" ~ "Male",
      group == "Female" ~ "Female"
    ),
    outcome = case_when(
      outcome == "proBEH" ~ "Problem Behavior",
      outcome == "socSKI" ~ "Social Skill"
    ),
    reporter = case_when(
      reporter == "adole" ~ "Adolescent",
      reporter == "careg" ~ "Caregiver"
    ),
    estimate = as.numeric(str_extract(boot, "^[\\d.-]+")),
    ll = as.numeric(str_extract(boot, "(?<=\\()[\\d.-]+")),
    ul = as.numeric(str_extract(boot, "(?<=, )[\\d.-]+")),
  ) %>%
  select(-boot)
```

#viz formatting
```{r}
library(tidyverse)
library(patchwork)
library(ggh4x)

# Label is only exposure_period
qgc_wt <- qgc_wt %>%
    mutate(label = exposure_period)

# Overall estimate + CI
est_df <- qgc_wt %>%
    select(label, outcome, reporter, group, estimate, ll, ul) %>%
    mutate(facet = "Overall Estimate", OPE = "Overall")

# OPE weights
ope_df <- qgc_wt %>%
    rename(overall_est = estimate) %>%
    pivot_longer(cols = c(BCEP, BDCIPP, DNBP, DPHP),
                 names_to = "OPE", values_to = "estimate") %>%
    mutate(ll = NA_real_, ul = NA_real_, facet = "OPE Weights") %>%
    select(label, outcome, reporter, group, estimate, ll, ul, OPE, facet)

# Combine
plot_df <- bind_rows(est_df, ope_df) %>%
    mutate(facet = factor(facet, levels = c("Overall Estimate", "OPE Weights")))

# Fixed axis limits for alignment
est_lims <- plot_df %>% filter(facet == "Overall Estimate") %>%
    summarise(lo = min(ll, na.rm = TRUE) - 0.5, hi = max(ul, na.rm = TRUE) + 0.5)
ope_lims <- plot_df %>% filter(facet == "OPE Weights") %>%
    summarise(lo = min(estimate, na.rm = TRUE) - 0.1, hi = max(estimate, na.rm = TRUE) + 0.1)

# Helper functions
make_p1 <- function(data) {
    ggplot(data, aes(x = estimate, y = label)) +
        geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
        geom_linerange(aes(xmin = ll, xmax = ul), na.rm = TRUE) +
        geom_point(size = 3) +
        scale_x_continuous(limits = c(est_lims$lo, est_lims$hi)) +
        facet_nested(group + reporter ~ ., switch = "y",
                     scales = "free_y", space = "free_y") +
        labs(x = NULL, y = NULL) +
        theme_minimal() +
        theme(
            text = element_text(color = "black"),
            axis.text = element_text(color = "black", size = 10),
            strip.placement = "outside",
            strip.text.y.left = element_text(face = "bold", size = 9, angle = 0, color = "black"),
            strip.background = element_blank(),
            panel.spacing.y = unit(0.3, "lines"),
            panel.grid.minor = element_blank()
        )
}

make_p2 <- function(data) {
    ggplot(data, aes(x = estimate, y = label)) +
        geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
        geom_point(aes(color = OPE, shape = OPE), size = 3) +
        scale_color_manual(values = c("BCEP" = "#E41A1C", "BDCIPP" = "#377EB8",
                                      "DNBP" = "#4DAF4A", "DPHP" = "#FF7F00")) +
        scale_shape_manual(values = c("BCEP" = 16, "BDCIPP" = 17,
                                      "DNBP" = 15, "DPHP" = 18)) +
        scale_x_continuous(limits = c(ope_lims$lo, ope_lims$hi)) +
        facet_nested(group + reporter ~ .,
                     scales = "free_y", space = "free_y") +
        labs(x = NULL, y = NULL,
             color = "OPE Biomarkers", shape = "OPE Biomarkers") +
        theme_minimal() +
        theme(
            text = element_text(color = "black"),
            axis.text.x = element_text(color = "black", size = 10),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            strip.text.y = element_text(color = "transparent"),
            legend.position = "bottom",
            legend.justification = "left",
            legend.text = element_text(color = "black"),
            legend.title = element_text(color = "black"),
            panel.spacing.y = unit(0.3, "lines"),
            panel.grid.minor = element_blank(),
            strip.background = element_blank()
        )
}

# Build rows split by outcome
outcomes <- unique(plot_df$outcome)

rows <- lapply(seq_along(outcomes), function(i) {
    o <- outcomes[i]
    d1 <- plot_df %>% filter(facet == "Overall Estimate", outcome == o)
    d2 <- plot_df %>% filter(facet == "OPE Weights", outcome == o)
    
    label <- ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = o,
                 fontface = "bold", size = 4, angle = 90, color = "black") +
        theme_void()
    
    p1_title <- if (i == 1) "Bootstrap Estimate" else NULL
    p2_title <- if (i == 1) "OPE Biomarker Weights" else NULL
    
    left <- make_p1(d1) + labs(title = p1_title) +
        theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5, color = "black"))
    
    right <- make_p2(d2) + labs(title = p2_title) +
        theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5, color = "black"))
    
    label + left + right + plot_layout(widths = c(0.3, 1.5, 3))
})

# Stack and collect legend
rows[[1]] / rows[[2]] +
    plot_layout(heights = c(1, 1), guides = "collect") &
    theme(legend.position = "bottom")

# Save
ggsave("~/Documents/ope_ssis/res_sing_imput/qgc/qgc_estimate_weights.tiff",
       width = 10, height = 8.5,
       dpi = 300, bg = "white")
```




