---
title: "099_misc"
author: "Puvvula"
date: "2023-07-19"
output: pdf_document
---

#to generate rda file for exposure-outcome combination by imputed set

#Model exposure-response associations using GEE model - to save each model to rda object
```{r}
dat<- read_csv(paste0(dat, "ope_ssis_imputed.csv"))

model<- geeglm(ss_std_score_c ~ log(dphp)*visit + 
                 gender + child_race + mat_age + mari_st_p4 + mom_edu_p4_cat+
                 mid_income_p4 + cotinine, 
  data = imp_dat_bcep, 
  id = id, 
  family="gaussian", corstr="independence")
```

```{r}
runGeeglm <- function(data, output_folder, exposures, outcomes, covariates, imputation, p_id) {
  # Create the output folder if it doesn't exist
  if (!dir.exists(output_folder)) {
    dir.create(output_folder)
  }

  # Loop through each combination of exposures, outcomes, and imputation levels
  for (exposure in exposures) {
    for (outcome in outcomes) {
      for (imp_level in unique(imputation)) {
        # Prepare the file name for the output
        prefix <- paste0("imp", imp_level)
        file_name <- paste0(output_folder, "/", prefix, "_", exposure, "_", outcome, ".rda")

        # Set the response variable
        response <- data[[outcome]]

        # Set the predictor variables
        predictors <- c(paste0("log(", exposure, ")*visit"), covariates)
        formula <- as.formula(paste(outcome, "~", paste(predictors, collapse = " + ")))

        # Subset the data for the specific imputation level
        imp_data <- data[imputation == imp_level, ]

        # Fit the geeglm model
        model <- geeglm(formula, data = imp_data, id = p_id, family = "gaussian", corstr = "independence")

        # Save the model object to an RDA file with the modified file name
        save(model, file = file_name)

        # Print a message indicating the completion of the model
        cat("Geeglm model saved for imputation level", imp_level, ",", exposure, "and", outcome, "\n")
      }
    }
  }
}
```

#testing
```{r}
runGeeglm(datasets = c("imp_dat_bcep", "imp_dat_bdcipp", "imp_dat_dnbp", "imp_dat_dphp"), 
          covariates = c("gender", "child_race", "mat_age", "mari_st_p4", "mom_edu_p4_cat",
                         "mid_income_p4", "cotinine" , "bdi_p4", "lead", "rfr_gs_t"),
          outcome = c("ss_std_score_c", "com_raw_c", "coop_raw_c", "assert_raw_c","res_raw_c",
                      "emp_raw_c", "eng_raw_c", "self_raw_c", "pb_std_score_c", "ext_raw_c",
                      "bul_raw_c", "hyp_raw_c", "int_raw_c", "ss_std_score_p", "com_raw_p",
                      "coop_raw_p", "assert_raw_p", "res_raw_p", "emp_raw_p", "eng_raw_p",
                      "self_raw_p", "pb_std_score_p", "ext_raw_p", "bul_raw_p", "hyp_raw_p",
                      "int_raw_p", "as_raw_p"),
          p_id="participant_id",
          outputFolder = "~/Documents/ope_ssis/gee_mi_res_feb/")

```

#function to compile results into a file
```{r}
compile_gee_mi_results <- function(folder_location) {
  # List all .rda files in the given folder
  rda_files <- list.files(path = folder_location, pattern = "\\.rda$", full.names = TRUE)
  
  # Create an empty tibble to store the combined results
  combined_results <- tibble()
  
  # Loop through each .rda file
  for (file in rda_files) {
    # Load the .rda file
    load(file)
    
    # Extract tidy results with confidence intervals
    tidy_results <- tidy(model, conf.int = TRUE)
    
    # Add the object name as a new column in the tidy results
    tidy_results <- tidy_results %>% mutate(file_name = basename(file))
    
    # Append the tidy results to the combined_results tibble
    combined_results <- bind_rows(combined_results, tidy_results)
    
    # Remove the loaded object from the environment
    rm(model)
  }
  
  # Return the combined results
  return(combined_results)
}
```

#get results
```{r}
mi_result <- compile_gee_mi_results("~/Documents/ope_ssis/result/gee_mi_res")

write_csv(mi_result, paste0(res, "mi_res_estimates.csv"))
```


#function to run GEE - using dataframes from MICE miltiple imputation
```{r}
runGeeglm <- function(datasets, outcomes, outputFolder) {
  for (dataset in datasets) {
    # Extract the last part of the input data name
    exposure <- tail(strsplit(dataset, "_")[[1]], 1)
  
    # Load the dataset
    data <- get(dataset)
  
    # Iterate over each outcome
    for (outcome in outcomes) {
      # Perform geeglm analysis
      fit <- with(data, geeglm(get(outcome) ~ log(get(exposure))*visit + 
                                gender + child_race + mat_age + mari_st_p4 + mom_edu_p4_cat +
                                mid_income_p4 + cotinine, 
                              id = participant_id, 
                              family = "gaussian", corstr = "independence"))
    
      # Save the output to the provided folder in RDA format
      outputName <- paste(outputFolder, "/", dataset, "_", outcome, "_", exposure, ".rda", sep = "")
      save(fit, file = outputName)
    }
  }
}
```

#for testing only
```{r}
runGeeglm(datasets = c("imp_dat_dphp","imp_dat_bcep", "imp_dat_bdcipp",  "imp_dat_dnbp"),
                  outcomes = c("ss_std_score_c", "com_raw_c", "coop_raw_c", "assert_raw_c","res_raw_c",
                              "emp_raw_c", "eng_raw_c", "self_raw_c", "pb_std_score_c", "ext_raw_c",
                              "bul_raw_c", "hyp_raw_c", "int_raw_c", "ss_std_score_p", "com_raw_p",
                              "coop_raw_p", "assert_raw_p", "res_raw_p", "emp_raw_p", "eng_raw_p",
                              "self_raw_p", "pb_std_score_p", "ext_raw_p", "bul_raw_p", "hyp_raw_p",
                              "int_raw_p", "as_raw_p"),
                  outputFolder = "~/Documents/ope_ssis/result/gee_mi_res_feb/")
```

#function to extract GEE estimates from rda objects
```{r}
extractGEE <- function(input_folder, output_folder) {
  # Get a list of RDA files in the input folder
  rda_files <- list.files(input_folder, pattern = "\\.rda$", full.names = TRUE)
  
  # Initialize an empty dataframe to store the results
  data <- data.frame()
  
  # Loop through each RDA file
  for (file in rda_files) {
    # Load the RDA file
    load(file)
    
    # Perform operations and extract required rows
    extracted_data <- as_tibble(tidy(pool(fit), conf.int = TRUE)) %>%
      slice(2, 12, 13)
    
    # Get the filename without path or extension
    filename <- str_remove(basename(file), "\\.[^.]+$")
    
    # Trim the filename to remove the "imp_dat_" prefix
    trimmed_filename <- str_remove(filename, "^imp_dat_")
    
    # Split the filename into exposure and outcome variables
    split_strings <- str_split(trimmed_filename, "_")
    exposure <- split_strings[[1]][1]
    outcome <- paste(split_strings[[1]][2:(length(split_strings[[1]]) - 1)], collapse = "_")
    
    # Add the exposure and outcome variables
    extracted_data$exposure <- exposure
    extracted_data$outcome <- outcome
    
    # Bind the extracted data to the existing dataframe
    data <- bind_rows(data, extracted_data)
  }
  
  # Export the dataframe as a CSV to the output folder
  output_file <- file.path(output_folder, "output.csv")
  write_csv(data, output_file)
}
```

#linear contribution
```{r}
load("~/Documents/ope_ssis/result/gee_mi_res_march/imp_dat_bcep_as_raw_p_bcep.rda")

m_1<-fit$analyses[1][[1]]

m_1$formula <- as.formula(gsub("log\\(get\\(exposure\\)\\)", "bcep", deparse(m_1$formula)))
names(m_1$coefficients) <- gsub("log\\(get\\(exposure\\)\\)", "bcep", names(m_1$coefficients))

confint(
  glht(
    m_1, 
    linfct = c("bcep=0",
               "bcep + bcep:visitprenatal_26W=0",
               "bcep + bcep:visitdelivery=0",
               "bcep + bcep:visitpostnatal_1_yr=0",
               "bcep + bcep:visitpostnatal_2_yr=0",
               "bcep + bcep:visitpostnatal_3_yr=0",
               "bcep + bcep:visitpostnatal_5_yr=0",
               "bcep + bcep:visitpostnatal_8_yr=0")
  )
)
```

```{r}

```


